A Zero-Knowledge Protocol for Keystroke Authentication

### 2014-12-15 paper/paper.tex
basic protocol data

\newpage
In our implementation, we modified the fuzzy extraction principle in order to hash each keystroke in a locally sensitive way. This is because we want to ignore small fluctuations and noises in between strokes.  Our implementation involves picking a error tolerance parameter $\delta$ such that similar keystroke timings will be viewed indiscriminately to the hashing algorithm.  Adjusting $\delta$ allows us to control how much error we want to allow per keystroke. After a series of algebraic calculations, we compare the exponentiated extracted hashed values to the exponentiated extracted hash values stored in the server.

\section{Results}
\label{sec:results}
We evaluate performance based on two main metrics: precision (how easily a person can repeat their own typing pattern), and protection (how easily an imposter can forget the typing pattern). As stated in Section~\ref{sec:bg-keystroke}, this adversary may not be an actual person; it may be a bot that is programmed to elicit the keystroke timing.  We test the performance of our application with four participants, all with varying typing speeds, preferred hand layouts, and dominant hands.  We measured $\epsilon^{'}$ which is the weighted difference in average typing speed the target user was typing at, compared to our threshold level of $\epsilon=40$. We then measured how consistently the target user could replicate that result, and measured that as a percentage. We then allowed the three other participants to attempt to forge the typing signature, having not seen the target user's test trials.

\begin{table}[h!]
\centering
\begin{tabular}{||c c c c||} 
 \hline
 Length of Name & $\epsilon^{'}$ & Success by User (\%) & Success by Imposter (\%) \\ [0.5ex] 
 \hline\hline
 3 & 112 & 4 & 0 \\ 
 5 & 59 & 10 & 5 \\
 8 & 47 & 85 & 23 \\ 
 11 & 34 & 87 & 43 \\
 13 & 32 & 87 & 35 \\ [1ex] 
 \hline
\end{tabular}
\caption{Keystroke variation by length of name.}
\label{table:1}
\end{table}

Table \ref{table:1} documents repeated test trials with $\delta=0.05$.  Trials were conducted between 20 and 40 times with varying word lengths.  We can see the trend of successful authentications with increasing word length.  This is most likely attributed to the fact that typing patterns are more distinct with longer pieces of text.  On the flip side, imposters are also more likely to successfully impersonate the target user with a longer text, since our algorithm does not differentiate where the difference in time accumulated; in other words, if an imposter's speed did not match in the beginning, (s)he can make up for it by matching in the end.  The error would be spread across the length of the text and therefore validate an imposter.

>>>>

\newpage
In our implementation, we modified the fuzzy extraction principle in order to hash each keystroke in a locally sensitive way. This is because we want to ignore small fluctuations and noises in between strokes.  Our implementation involves picking a error tolerance parameter $\delta$ such that similar keystroke timings will be viewed indiscriminately by the hashing algorithm.  Adjusting $\delta$ allows us to control how much error we want to allow per keystroke. After a series of algebraic calculations, we compare the exponentiated extracted hashed values to the exponentiated extracted hash values stored in the server.

\section{Results}
\label{sec:results}
We evaluate performance based on two main metrics: precision (how easily a person can repeat their own typing pattern), and protection (how easily an imposter can forget the typing pattern). As stated in Section~\ref{sec:bg-keystroke}, this adversary may not be an actual person; it may be a bot that is programmed to elicit the keystroke timing.  We test the performance of our application with four participants, all with varying typing speeds, preferred hand layouts, and dominant hands.  We measured $\epsilon^{'}$ which is the weighted difference in average typing speed the target user was typing at, compared to our threshold level of $\epsilon=40$ in the zero-knowledge case and $\epsilon=50$ for the basic protocol. We then measured how consistently the target user could replicate that result, and measured that as a percentage. We then allowed the three other participants to attempt to forge the typing signature, having not seen the target user's test trials.  Identical words were selected  as the control for both protocols in order to measure the variance between the two success rates.

\begin{table}[h!]
\centering
\begin{tabular}{||c c c c||} 
 \hline
 Length of Name & $\epsilon^{'}$ & Success by User (\%) & Success by Imposter (\%) \\ [0.5ex] 
 \hline\hline
 3 & 14 & 100 & 100 \\ 
 5 & 16 & 87 & 65 \\
 8 & 47 & 85 & 25 \\ 
 11 & 43 & 87 & 12 \\
 13 & 47 & 65 & 0 \\ [1ex] 
 \hline
\end{tabular}
\caption{Basic Protocol: Keystroke variation by length of name.}
\label{table:1}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{||c c c c||} 
 \hline
 Length of Name & $\epsilon^{'}$ & Success by User (\%) & Success by Imposter (\%) \\ [0.5ex] 
 \hline\hline
 3 & 112 & 4 & 0 \\ 
 5 & 59 & 10 & 5 \\
 8 & 47 & 85 & 23 \\ 
 11 & 34 & 87 & 43 \\
 13 & 32 & 87 & 35 \\ [1ex] 
 \hline
\end{tabular}
\caption{Zero-Knowledge Protocol: Keystroke variation by length of name.}
\label{table:2}
\end{table}

Table \ref{table:1} is the implementation of the Basic Protocol, as outlined in the previous section.  Trials were conducted between 20 and 40 times with varying word lengths.  The value of the timings were compared directly, allowing an error of $\epsilon=50$. The $\epsilon^{'}$ in \ref{table:1} indicates the average error value the target user was scoring.   From these preliminary results, we note some important trends; as word length increases, typing signatures become more unique and therefore, adversary forgery success rate drops significantly.  However, this also affects the target user's ability to recreate his or her typing signature consistently.  There appears to be a tradeoff between security and convenience.

Table \ref{table:2} documents repeated test trials with $\delta=0.05$.  Trials were conducted between 20 and 40 times with varying word lengths.  We can see the trend of successful authentications with increasing word length.  This is most likely attributed to the fact that typing patterns are more distinct with longer pieces of text.  On the flip side, imposters are also more likely to successfully impersonate the target user with a longer text, since our algorithm does not differentiate where the difference in time accumulated; in other words, if an imposter's speed did not match in the beginning, (s)he can make up for it by matching in the end.  The error would be spread across the length of the text and therefore validate an imposter.
